\input{preamble.tex}
\begin{document}
\section{Small Screens Optimisation}

\subsection{Constrains and Limitations}
On small screens horizontal scroll bar is not an option, since it breaks all the dynamics of the user experience, the processor is slower, the memory is limited so loading large file sizes is not a plus.\\
\\
\noindent \textbf{For small screen some limitations should be considered and applied to the design:}\\
- One column layout only, this will avoid the horizontal scroll bar.\\
- HTML optimised using efficient tags and attributes.\\
- CSS optimised with efficient styles proprieties.\\
- Minimise the amount of decorative images, or if needed, use of pre loaders (scripts that only show the page when fully loaded) so the user experience is not affected.\\
- Write good alternative text (attribute 'alt')  for images, so in case the image is not needed an alternative text could be shown for example navigation icons.\\
- Avoid effects that need mouse or keyboard events.\\
- Avoid the use of Javascript on layout design since some mobile browsers may be set to block this scripts.

\subsubsection{Screen Sizes Proposal}
Smaller screens will always be the tablets PCs or the ones that fits on pockets like smartphones.\\
Keeping this in mind the most used screen sizes are:\\
- 320x480 Screen Size ( eg. iPhone )\\ 	% Tested on: iPhone3G(OK),iPhone 4(OK)
- 480x800 Screen Size (eg. HTC Titan, HTC Desire)\\ %HTC desire(NOK)
- 1024x600 Screen Size ( eg. Samsung Galaxy )\\
- 1024x768 Screen Size ( eg. iPad)\\
\\
\noindent \textbf{List of the most popular smartphones and tablets}\\
- Samsung Galaxy Player 5.0 (released 2011)\\
- Samsung Galaxy S II (released 2011)\\
- Apple iPhone 4S (released 2011)\\
- Apple iPhone 4 (released 2010)\\
- HTC Titan (released 2011)\\

\subsection{Page Layout}

For a long time tables were used to make the webpages layout, this was made by nesting HTML table tags to create for example multi column pages, separate headers, main and footers, etc..

As solution, HTML should be used to the intellectual content of the page, while CSS determine how to represent the content to the user. This method of design improves the time of development since theres no more nested tables the HTML tags become more understandable, so create and redesigned a webpage becomes more easy and less time consuming, improves search engines rankings and files sizes diminished noticeable.\\

\subsubsection{Avoiding Images}

All images are loaded to the memory, being this very limited on smartphones and tablet computers, its necessary to "hide" some images so the browser won't load them, this have to be done without compromising the user experience.

If images are needed they should be scaled down on a image editor (e.g. PhotoShop), this way the size of the image ( file size ) will get smaller so it will take less memory space and less time to be loaded and shown to the user.

User should always be informed when the page is not completely loaded, this can be done using a preload script (script that hides the content of a page showing a loading bar until the page finish loading). This kind of scripts avoids a bad navigation experience to the user.

\subsubsection{Re-Design}

To make a page suitable for a small screen, the layout should be re-designed, in this case, only the CSS will be changed and the HTML will be optimised for mobile web browsers.

When the webpage is open, a CSS ( Cascade Style Sheet ) is loaded defining the layout that correspond to the screen size.

\subsubsection{Handling CSS Media Queries}  % CSS

CSS is able to get information about almost all newer smartphone devices, information such as orientation and screen size is some information that is retrieved by the browser.\\
\begin{lstlisting}
	/* iPhone portrait*/
	@media only screen and (max-width:320px) and (orientation:portrait)
	 {
	 }
\end{lstlisting}
This line of CSS media queries will ensure that screens with a maximum width of 320px and orientation portrait will use the layout defining between brackets. Media queries are conditions that defines if blocks of CSS properties are going to be loaded or not. This will be a great help to re-design a website layout since different pages can be already set to link to this style sheet file. \\

\begin{figure}[!h]
	\begin{tabular}{| c | c | c |}
	\hline
	\textbf{Device} & \textbf{Orientation} & \textbf{Media Queries} \\ \hline
	iPhone & Portrait & @media only screen and (max-width:320px) and (orientation:portrait)\\ \hline
	iPhone & Landscape & @media only screen and (max-width:480px) and (orientation:landscape)\\ \hline
	HTC & Portrait & @media only screen and (max-width:480px) and (orientation:portrait)\\ \hline
	HTC & Landscape & @media only screen and (max-width:800px) and (orientation:landscape)\\ \hline
	iPad / Galaxy & Portrait & @media only screen and (max-width:600px) and (orientation:portrait)\\
	\hline
	\end{tabular}
\end{figure}

\subsubsection{User Experience}	    % CSS

\input{smallreferences.tex}
\end{document}